Competing Bandits: The Perils of Exploration under Competition

### 2018-09-05 paper/content/barriers.tex
stuff

To answer these questions, we isolate the effect from each of the two
advantages and run two additional experiments. In the first experiment
(called \emph{reputation erased experiment}), the reputation of the
incumbent is artificially reset upon the entry of the entrant {so that
  the incumbent's reputation score is only based on the rewards
  generated in the warm start period}. In the second experiment
(called \emph{information erased experiment}), the information gained
by the incumbent is erased upon the entry of the entrant in the sense
that its posterior is reset to the initial prior. The results of the
experiments are presented in Table~\ref{barrier_exp}.

\begin{table*}[h]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
\hline
  & \multicolumn{3}{c|}{Information erased}
  & \multicolumn{3}{c|}{Reputation erased}\\
\hline
& $\TS$  & $\DEG$  & $\DG$ 
& $\TS$  & $\DEG$  & $\DG$
\\\hline
$\TS$ 
    & \makecell{\textbf{0.021}$\pm$0.009}
    & \makecell{\textbf{0.16}$\pm$0.02} 
    & \makecell{\textbf{0.21} $\pm$0.02} 
    %%
    & \makecell{\textbf{0.0096}$\pm$0.006}
    & \makecell{\textbf{0.11}$\pm$0.02}
    & \makecell{\textbf{0.18}$\pm$0.02} 
    \\ \hline
$\DEG$ 
    & \makecell{\textbf{0.26}$\pm$0.03} 
    & \makecell{\textbf{0.3}$\pm$0.02} 
    & \makecell{\textbf{0.26}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.073}$\pm$0.01}
    & \makecell{\textbf{0.29}$\pm$0.02}
    & \makecell{\textbf{0.25}$\pm$0.02} 
    \\ \hline
$\DG$ 
    & \makecell{\textbf{0.34}$\pm$0.03} 
    & \makecell{\textbf{0.4}$\pm$0.03} 
    & \makecell{\textbf{0.33}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.15}$\pm$0.02}
    & \makecell{\textbf{0.39}$\pm$0.03}
    & \makecell{\textbf{0.33}$\pm$0.02}
    \\\hline
\end{tabular}
\caption{Information-erased vs. reputation-erased experiment, on Heavy-Tail MAB instance. Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}). The cell specifies the entrant's market share for the rounds in which hit was present: the average (in bold) and the 95\% confidence interval. NB: smaller average is better for the incumbent.}
\label{barrier_exp}
\end{table*}

@@

\begin{finding}
\swedit{During the rounds in which the entrant is present, the incumbent
  with either data or reputational advantage alone gets a larger share
  than the share \gaedit{in the permanent duopoly, regardless of the algorithm she plays}.  When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two
  barriers.}
\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
\end{finding}

\swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}

\gaedit{While the data and reputation advantages seem to serve as substitutes for \DEG and \DG, it is interesting that under \TS the data advantage is larger. The reason for this is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}. Doing purposeful exploration in competition may have immediate adverse reputational consequences because exploration sometimes involves picking bad arms. However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms. As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.

>>>>

To answer these questions, we isolate the effect from each of the two
advantages and run two additional experiments. In the first experiment
(called \emph{reputation erased experiment}), the reputation of the
incumbent is artificially reset upon the entry of the entrant {so that
  the incumbent's reputation score is only based on the rewards
  generated in the warm start period}. In the second experiment
(called \emph{information erased experiment}), the information gained
by the incumbent is erased upon the entry of the entrant in the sense
that its posterior is reset to the initial prior. The results of the
experiments are presented in Table~\ref{barrier_exp}, which lead to
our Finding~\ref{barrier-find}.

\begin{table*}[h*]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
\hline
  & \multicolumn{3}{c|}{Information erased}
  & \multicolumn{3}{c|}{Reputation erased}\\
\hline
& $\TS$  & $\DEG$  & $\DG$ 
& $\TS$  & $\DEG$  & $\DG$
\\\hline
$\TS$ 
    & \makecell{\textbf{0.021}$\pm$0.009}
    & \makecell{\textbf{0.16}$\pm$0.02} 
    & \makecell{\textbf{0.21} $\pm$0.02} 
    %%
    & \makecell{\textbf{0.0096}$\pm$0.006}
    & \makecell{\textbf{0.11}$\pm$0.02}
    & \makecell{\textbf{0.18}$\pm$0.02} 
    \\ \hline
$\DEG$ 
    & \makecell{\textbf{0.26}$\pm$0.03} 
    & \makecell{\textbf{0.3}$\pm$0.02} 
    & \makecell{\textbf{0.26}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.073}$\pm$0.01}
    & \makecell{\textbf{0.29}$\pm$0.02}
    & \makecell{\textbf{0.25}$\pm$0.02} 
    \\ \hline
$\DG$ 
    & \makecell{\textbf{0.34}$\pm$0.03} 
    & \makecell{\textbf{0.4}$\pm$0.03} 
    & \makecell{\textbf{0.33}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.15}$\pm$0.02}
    & \makecell{\textbf{0.39}$\pm$0.03}
    & \makecell{\textbf{0.33}$\pm$0.02}
    \\\hline
\end{tabular}
\caption{Information-erased vs. reputation-erased experiment, on Heavy-Tail MAB instance. Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}). The cell specifies the entrant's market share for the rounds in which hit was present: the average (in bold) and the 95\% confidence interval. NB: smaller average is better for the incumbent.}
\label{barrier_exp}
\end{table*}

@@

\begin{finding}\label{barrier-find}
\swedit{During the rounds in which the entrant is present, the incumbent
  with either data or reputational advantage alone gets a larger share
  than the share \gaedit{in the permanent duopoly, regardless of the algorithm she plays}.  When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two
  barriers.}
\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
\end{finding}

% \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}

\gaedit{While the data and reputation advantages seem to serve as substitutes for \DEG and \DG, it is interesting that under \TS the data advantage is larger. The reason for this is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}. In contrast, purposeful exploration in competition may have immediate adverse reputational consequences. However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms. As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.
